`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-20 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
added some clarification to the appendix and fixed some notation
general paper edits

The analytic derivation of the inner product of the \emph{n}-dimensional Gaussians is provided in~\ref{ap:InnerProdOfGaussians}. Equation~\ref{eq:ReducedForm} can be simplified by exploiting the symmetry (isotropy) of the connectivity kernel where

@@

which is constant $L \times n_{\theta}$ matrix that can be defined analytically (see~\ref{ap:ConvOfGaussians} for analytic convolution of two Gaussians), where $n_{\theta}$ is the number of connectivity basis functions. Now substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReducedForm} gives

@@

where $L$ is the number of basis functions used to decompose the neural field and 

@@

% \begin{equation}
% \mathbf{C} = \left[\begin{array}{{cccc}}
% \int_{\Omega}m(\mathbf{r}_1 - \mathbf{r}')\boldsymbol{\phi}_1(\mathbf{r}')d\mathbf{r}' & \int_{\Omega} m(\mathbf{r}_1 - \mathbf r')\boldsymbol \phi_2(\mathbf r')d\mathbf r' & \dots
% &\int_{\Omega}m(\mathbf r_1 - \mathbf r')\boldsymbol \phi_n(\mathbf
% r')d\mathbf r' \\
% \int_{\Omega}m(\mathbf r_2 - \mathbf r')\boldsymbol \phi_1(\mathbf
% r')d\mathbf r'&\int_{\Omega}m(\mathbf r_2 - \mathbf r')\boldsymbol
% \phi_2(\mathbf r')d\mathbf r'& \dots &\int_{\Omega}m(\mathbf r_2 -
% \mathbf r')\boldsymbol \phi_n(\mathbf r')d\mathbf r'\\
% \vdots&\vdots&\ddots&\vdots\\
% \int_{\Omega}m(\mathbf r_{n_y} - \mathbf r')\boldsymbol \phi_1(\mathbf
% r')d\mathbf r'&\int_{\Omega}m(\mathbf r_{n_y} - \mathbf r')\boldsymbol
% \phi_2(\mathbf r')d\mathbf r'& \dots &\int_{\Omega}m(\mathbf r_{n_y} -
% \mathbf r')\boldsymbol \phi_n(\mathbf r')d\mathbf r'
% \end{array}\right],
% \end{equation}
% and $m(\mathbf r_n - \mathbf{r}')$ is the observation kernel.
\section{Spectral Analysis and Model Selection}\label{SpectralAnalysisSection} Spectral analysis has been used to identify both the number of sensors and the number of basis functions required to reconstruct the neural field from sampled observations \cite{Sanner1992,Scerri2009}. Based on a two-dimensional extension of Shannon's sampling theorem \cite{Peterson1962}, the spatial bandwidth of the observed field can be used to provide a lower bound on both the number of sensors and the number of basis functions required to capture the dominant spectral characteristics of the neural field.

@@

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. This motivates the use of sensors with wider bandwidths (narrower in space). Nevertheless, this choice would require more sensors in order to satisfy Shannon's sampling theorem for a reasonable spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors used and the computational demands of the estimation procedure when designing experiments.

@@

Alternatively, given the basis function width, the spatial cut-off frequency of the field modelled by the decomposition can be determined from spectral analysis. This is particularly useful when considering the spatially distributed basis functions, since Shannon's sampling theorem is derived for perfect point sensors. To calculate the cut-off frequency imposed by the basis function, $\boldsymbol{\nu}_{c\phi}$, we rearrange equations~\ref{eq:WidthCutOffRelationship} and \ref{eq:WidthFrequencyRelationship} giving
\begin{equation}\label{eq:CutoffFromBasisFuncWidth}
	\boldsymbol{\nu}_{c\phi}=\frac{1}{\pi\sigma_{\phi}}\sqrt{\frac{\ln2}{2}}.
\end{equation}

@@

The unscented Rauch Tung Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is a scaling parameter (see~\cite{Haykin2001} for more details).

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, needs to be computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, instead of augmenting the state vector with the noise terms with the added benefit of reducing the dimension and the number of sigma points required in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This feature is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state vector sequence estimate. To create the least squares parameter estimator, we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the URTSS, we can write
\begin{eqnarray*}
	\mathbf x_{1} &=& \mathbf{q}(\mathbf x_0) \boldsymbol{\theta}+\xi\mathbf x_0+\mathbf e_0 \\
	\mathbf x_{2} &=& \mathbf{q}(\mathbf x_1) \boldsymbol{\theta}+\xi\mathbf x_1+\mathbf e_1  \\
	&\vdots& \\
	\mathbf x_{T}&=&\mathbf{q}(\mathbf x_{T-1}) \boldsymbol{\theta}+\xi\mathbf x_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\mathbf x_{1}\\
		\mathbf x_{2}\\
		\vdots\\
		\mathbf x_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\mathbf x_0)& \mathbf x_{0}\\
		\mathbf q(\mathbf x_1)& \mathbf x_{1}\\
		\vdots & \vdots\\
		\mathbf q(\mathbf x_{T-1})& \mathbf x_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates, $ \mathcal{\hat{W}}$, are
\begin{equation}
	\mathcal{\hat{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}
% The accuracy of the approximation is linked to the time step used in the temporal discretisation of the system, where the state transition appears more linear with a finer resolution. 
% Alternatives to the UKF are the Extended Kalman Filter (EKF)~\cite{Haykin2001} and the Sequential Monte Carlo (SMC) filter~\cite{doucet2001}. The EKF approximates the state transition equation by linearising about the current state estimates (by calculating the Jacobian) at each time instance. The linearisation maintains the Gaussianality of the model. The unscented Kalman filter approximates the posterior state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, while maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, since the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the EKF can be problematic. SMC filtering can theoretically provide an exact posterior state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to high computation demands.

@@

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Eleven iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:ParametersConvergence} shows the rate of convergence, confirming the parameters converged, with the changes in parameters falling to less than $10^{-4}$ after six iterations.
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{The mean of the absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:ParametersConvergence}
\end{figure}

@@

\begin{figure}
    \centering
% \subfigure[][]{
\includegraphics{./Graph/KernelEstimate.pdf}%}
% \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Cross-section of connectivity kernel estimate. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the $95~\%$ confidence interval is shown by the shaded grey region.}
\label{fig:KernelEstimates}
\end{figure}
The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over time) of the root mean square error (MRMSE) (over space). The RMSE of the field estimates over time for 100 realisations is shown in Figure~\ref{fig:RMSE}, with the $95~\%$ confidence interval. At each time point the error is consistently around 0.5, which approximately $7.1$~\% of the field voltage range, which is $(-3.29,3.77)$~mV. The error in the reconstructed field is mostly due to the loss of high spatial frequency information due to the basis function decomposition as illustrated in Figure~\ref{fig:FieldEstimate}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:FieldEstimate}(b) that the estimated field contains the basic structure of the true field (Figure~\ref{fig:FieldEstimate}(a)), but is missing the higher-frequency detail.
  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:RMSE}
   \end{figure}

@@

The development of an estimation framework also has theoretical implications. The the ability to estimate parameters will allow testing of hypotheses that have been generated in theoretical studies. This has the potential to validate neural field models, and in doing so, provide new insights that may lead to a greater understanding of macro-scale neurodynamics. For example, using the proposed model-based framework we can reconstruct a neural field from data and estimate parameters and check if they correspond to the theoretically derived parameters where certain solutions can exist. In addition, the parameters and the state-space can be explored during cognitive tasks and states of arousal such as during sleep and anaesthesia.

In this paper, assumption was made that the chosen kernel basis functions can exactly represent the actual connectivity kernel. In future work, we plan to relax this assumption. We hope to achieve this by using the actual kernel that cannot exactly be presented by the chosen finite number of basis functions. In addition, we have assumed the neural field is isotropic and homogeneous. It is generally accepted that the intracortical connectivity structure can be accurately modelled as being isotropic. However, the assumption of homogeneity is quite strong. Future work should be focussed on developing estimation algorithms where assumption of homogeneity can be relaxed. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite, distant dependant propagation times into the parametric neural field equations.

@@

This Appendix provides the derivation of the field basis function cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centred at the origin (see equation~\ref{eq:BasisFunctionAtOrigin}) yields

@@

 \lefteqn{\boldsymbol\Phi(\boldsymbol \nu)=\int_{\mathcal R^n}{\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r}} \nonumber \\
&=&\int_{\mathcal R^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \nonumber \\

@@

\int_{\mathcal R^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r&& \nonumber \\

@@

and hence completing the derivation. For 3 dB attenuation at $\boldsymbol\nu_c$ we need to set

@@

 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2

@@

solving for $\sigma_{\nu}^2$ we get

@@

\Gamma_{i,j} = \phi_i\otimes\phi_j=\int_{\Omega}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf{r}.

@@

\phi_i\otimes\phi_j&=&\int_{\Omega}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&\int_{\Omega}\phi_i(2\boldsymbol \mu_i-\mathbf r)\phi_j(\mathbf r)d\mathbf r\\

>>>>

An analytic derivation of the inner product of \emph{n}-dimensional Gaussians is provided in~\ref{ap:InnerProdOfGaussians}. Equation~\ref{eq:ReducedForm} can be simplified by exploiting the symmetry (isotropy) of the connectivity kernel where

@@

which is a constant $L \times n_{\theta}$ matrix that can is defined analytically (see~\ref{ap:ConvOfGaussians} for analytic convolution of two Gaussians), where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity basis functions. Now substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReducedForm} gives

@@

@@

% \begin{equation}
% \mathbf{C} = \left[\begin{array}{{cccc}}
% \int_{\Omega}m(\mathbf{r}_1 - \mathbf{r}')\boldsymbol{\phi}_1(\mathbf{r}')d\mathbf{r}' & \int_{\Omega} m(\mathbf{r}_1 - \mathbf r')\boldsymbol \phi_2(\mathbf r')d\mathbf r' & \dots
% &\int_{\Omega}m(\mathbf r_1 - \mathbf r')\boldsymbol \phi_n(\mathbf
% r')d\mathbf r' \\
% \int_{\Omega}m(\mathbf r_2 - \mathbf r')\boldsymbol \phi_1(\mathbf
% r')d\mathbf r'&\int_{\Omega}m(\mathbf r_2 - \mathbf r')\boldsymbol
% \phi_2(\mathbf r')d\mathbf r'& \dots &\int_{\Omega}m(\mathbf r_2 -
% \mathbf r')\boldsymbol \phi_n(\mathbf r')d\mathbf r'\\
% \vdots&\vdots&\ddots&\vdots\\
% \int_{\Omega}m(\mathbf r_{n_y} - \mathbf r')\boldsymbol \phi_1(\mathbf
% r')d\mathbf r'&\int_{\Omega}m(\mathbf r_{n_y} - \mathbf r')\boldsymbol
% \phi_2(\mathbf r')d\mathbf r'& \dots &\int_{\Omega}m(\mathbf r_{n_y} -
% \mathbf r')\boldsymbol \phi_n(\mathbf r')d\mathbf r'
% \end{array}\right],
% \end{equation}
% and $m(\mathbf r_n - \mathbf{r}')$ is the observation kernel.
\section{Spectral Analysis and Model Selection}\label{SpectralAnalysisSection} Spectral analysis has been used to identify both the number of sensors and the number of basis functions required to reconstruct the neural field from sampled observations~\cite{Sanner1992,Scerri2009}. Based on a two-dimensional extension of Shannon's sampling theorem \cite{Peterson1962}, the spatial bandwidth of the observed field can be used to provide a lower bound on both the number of sensors and the number of basis functions required to capture the dominant spectral characteristics of the neural field.

@@

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. This motivates the use of sensors with wider bandwidths (narrower in space). Nevertheless, this choice would require more sensors in order to satisfy Shannon's sampling theorem for a given spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors used and the computational demands of the estimation procedure when designing experiments.

@@

Alternatively, given the basis function width, the spatial cut-off frequency of the field modelled by the decomposition can be determined from spectral analysis. This is particularly useful when considering the spatially distributed basis functions, since Shannon's sampling theorem is derived for perfect point sensors. To calculate the cut-off frequency imposed by the basis functions, $\boldsymbol{\nu}_{c\phi}$, we rearrange equations~\ref{eq:WidthCutOffRelationship} and \ref{eq:WidthFrequencyRelationship} giving
\begin{equation}\label{eq:CutoffFromBasisFuncWidth}
	\boldsymbol{\nu}_{c\phi}=\frac{1}{\pi\sigma_{\phi}}\sqrt{\frac{\ln2}{2}}.
\end{equation}

@@

The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, is also computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, instead of augmenting the state vector with the noise terms with the added benefit of reducing the dimension and the number of sigma points required in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This feature is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from an iteration of the URTSS, we can write
\begin{eqnarray*}
	\mathbf x_{1} &=& \mathbf{q}(\mathbf x_0) \boldsymbol{\theta}+\xi\mathbf x_0+\mathbf e_0 \\
	\mathbf x_{2} &=& \mathbf{q}(\mathbf x_1) \boldsymbol{\theta}+\xi\mathbf x_1+\mathbf e_1  \\
	&\vdots& \\
	\mathbf x_{T}&=&\mathbf{q}(\mathbf x_{T-1}) \boldsymbol{\theta}+\xi\mathbf x_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\mathbf x_{1}\\
		\mathbf x_{2}\\
		\vdots\\
		\mathbf x_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\mathbf x_0)& \mathbf x_{0}\\
		\mathbf q(\mathbf x_1)& \mathbf x_{1}\\
		\vdots & \vdots\\
		\mathbf q(\mathbf x_{T-1})& \mathbf x_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates, $ \mathcal{\hat{W}}$, are
\begin{equation}
	\mathcal{\hat{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}
% The accuracy of the approximation is linked to the time step used in the temporal discretisation of the system, where the state transition appears more linear with a finer resolution. 
% Alternatives to the UKF are the Extended Kalman Filter (EKF)~\cite{Haykin2001} and the Sequential Monte Carlo (SMC) filter~\cite{doucet2001}. The EKF approximates the state transition equation by linearising about the current state estimates (by calculating the Jacobian) at each time instance. The linearisation maintains the Gaussianality of the model. The unscented Kalman filter approximates the posterior state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, while maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, since the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the EKF can be problematic. SMC filtering can theoretically provide an exact posterior state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to high computation demands.

@@

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the state and parameter estimation algorithm, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Eleven iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:ParametersConvergence} shows the rate of convergence, confirming the parameters converged, with the changes in parameters falling to less than $10^{-4}$ after six iterations.
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{The mean of the absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:ParametersConvergence}
\end{figure}

@@

\begin{figure}
    \centering
% \subfigure[][]{
\includegraphics{./Graph/KernelEstimate.pdf}%}
% \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Cross-section of connectivity kernel estimate. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the $95~\%$ confidence interval is shown by the shaded grey region.}
\label{fig:KernelEstimates}
\end{figure}
The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) (over space). The RMSE of the field estimates over time for 100 realisations is shown in Figure~\ref{fig:RMSE}, with the $95~\%$ confidence interval. At each time point the error is consistently around 0.5, which approximately $7.1$~\% of the field voltage range, which is $(-3.29,3.77)$~mV. The error in the reconstructed field is mostly due to the loss of high spatial frequency information due to the basis function decomposition as illustrated in Figure~\ref{fig:FieldEstimate}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:FieldEstimate}(b) that the estimated field contains the basic structure of the true field (Figure~\ref{fig:FieldEstimate}(a)), but is missing the higher-frequency detail.
  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:RMSE}
   \end{figure}

@@

The development of an estimation framework also has theoretical implications. The the ability to estimate parameters will allow testing of hypotheses that have been generated in theoretical studies. This has the potential to validate neural field models, and in doing so, provide new insights that may lead to a greater understanding of macro-scale neurodynamics. For example, using the proposed model-based framework we can reconstruct a neural field from data and estimate parameters and check if they correspond to theoretically derived parameters where certain solutions can exist. In addition, the parameters and the state-space can be explored during cognitive tasks and states of arousal such as during sleep and anaesthesia.

In this paper, an assumption was made that the chosen kernel basis functions can exactly represent the actual connectivity kernel. In future work, we plan to relax this assumption. We hope to achieve this by using the actual kernel that cannot exactly be presented by the chosen finite number of basis functions. In addition, we have assumed the neural field is isotropic and homogeneous. It is generally accepted that the intracortical connectivity structure can be accurately modelled as being isotropic. However, the assumption of homogeneity is quite strong. Future work should be focussed on developing estimation algorithms where assumption of homogeneity can be relaxed. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite, and distant dependant propagation times into the parametric neural field equations.

@@

This Appendix provides the derivation of the field basis function cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centred at the origin (defined in equation~\ref{eq:BasisFunctionAtOrigin}) yields

@@

 \lefteqn{\boldsymbol\Phi(\boldsymbol \nu)=\int_{\mathbb{R}^n}{\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r}} \nonumber \\
&=&\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \nonumber \\

@@

\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r&& \nonumber \\

@@

For $3$~dB attenuation at $\boldsymbol\nu_c$ we need to set

@@

 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,

@@

since the maximum of Gaussian described in equation~\ref{eq:FTofGaussian} is at the origin. Now solving for $\sigma_{\nu}^2$ we get

@@

\Gamma_{i,j} = \phi_i\otimes\phi_j=\int_{\mathbb{R}^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf{r}.

@@

\phi_i\otimes\phi_j&=&\int_{\mathbb{R}^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&\int_{\mathbb{R}^n}\phi_i(2\boldsymbol \mu_i-\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
